Plant-pollintor networks worldwide are composed by the same specific building

### 2021-08-03 Draft/Draft.Rmd
Update Draft.Rmd

Following Simmons et al. [@simmons2019; @simmons2020], we broke down the plant-pollinator networks into their constituent motifs. Prior to analyses, we selected only robust links, that is, interactions whose frequency was greater than one (64.98 % interactions); and, then, we turned the quantitative networks into qualitative (or binary) ones, where interactions are present or absent. [Here I would present maybe full analysis, and without singletones in sup mat, especially if they are consistent]()

@@

Next, we calculated which functional groups were over or under-represented in different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. We extracted the position frequencies of all motifs from three to five nodes for each network and species by using the ‘bmotif’ package [@bmotifsimmons2019]. To estimate the position frequencies of each functional group in a given network, we added the frequencies of those species that belong to the group, and then, we normalised the resulting frequencies by dividing the position measure for each group by the total number of times that a group appears in any position within the same motif size class. Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did motif frequencies. To outline the general patterns of position frequencies across networks and functional groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the functional group identifier as an explanatory variable and the study identifiers as a random intercept. By adding the functional group estimates to the model intercept, we assessed the average motif frequency, after controlling the variation at the study level. To fit the LMMs models in our analyses, we used the ‘lmer’ package [@bates2015]. Finally, we visualized with the help of the package ComplexHeatmap version 2.6.2 [@gu2016] over- and under- representation of plant and pollinator functional groups on the different motif positions.

[potential additions: 
Finally,for each motif, we recover the number of times a given combination of functional groups emerges ...]

>>>>

Following Simmons et al. [@simmons2019; @simmons2020], we broke down the plant-pollinator networks into their constituent motifs. Prior to analyses, we turned the quantitative networks into qualitative (or binary) ones, where interactions are present or absent.

@@

Next, we calculated which functional groups were over or under-represented in different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. We extracted the position frequencies of all motifs from three to five nodes for each network and species by using the ‘bmotif’ package [@bmotifsimmons2019]. To estimate the position frequencies of each functional group in a given network, we added the frequencies of those species that belong to the group, and then, we normalised the resulting frequencies by dividing the position measure for each group by the total number of times that a group appears in any position within the same motif size class. Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did motif frequencies. To outline the general patterns of position frequencies across networks and functional groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the functional group identifier as an explanatory variable and the study identifiers as a random intercept. By adding the functional group estimates to the model intercept, we assessed the average motif frequency, after controlling the variation at the study level. Finally, we visualized with the help of the package ComplexHeatmap version 2.6.2 [@gu2016] over- and under- representation of plant and pollinator functional groups on the different motif positions.

To test the effect of singletones on the frequencies of motif and those of functional groups' positions, in Appendices XXX, we applied the previous analyzes to the robust versions of the networks in table S1, namely: networks that, prior to binarization, only contained interactions whose frequency was greater than one (64.98 % interactions).

Finally, we studied which motif combinations of functional groups (up to five nodes) are over or under-represented in 95% of our empirical networks (57 out of 60 arrangements, due to computational limitations to extract all motif combinations in the three networks with the highest number of links). To do so, for each of the 53,250 possible motif combinations, we estimated the observed and the expected probability of finding that combination in empirical networks, respectively. Then, we determined whether the observed probabilities are likely to come from the expected probabilities or not. To calculate the observed probability of a given combination $i$, $p_i^O$, we divided the number of times $i$ appears in our set empirical networks ($n_i^O$) by the sum of the number of times each possible combination appeared ($\sum_{k=1}^{53250} n_k^O$): $p_i^O = n_i^O / \sum_{k=1}^{53250} n_k^O$. Consequently, $\sum_{k=1}^{53250} p_k^O = 1$. To estimate the expected probability of a given motif combination $i$, $p_i^E$, firstly, we calculated the probability of finding a given functional group $x$ in the position $\alpha$ of $i$ ($p_i(x,\alpha)$). Then, by assuming the independency of $p_i(x,\alpha)$, we calculated the expected probability of the combination $i$ is equal to the product of the pairs $(x,\alpha)$ in $i$: $p_i^E = \prod _{(x,\alpha )\in i} p_i(x,\alpha)$. To obtain $p_i(x,\alpha)$, we calculated the number of times (or absolute frequency) that the functional group $x$ appears in the position $\alpha$ in 95% of our empirical networks, denoted as $n_i(x,\alpha)$. We controlled the variation at the network level and assessed the average value of $n_i(x,\alpha)$ by fitting a LMM per motif position, where the response variable is the number of times that a given position is observed per network, the explanatory variable is functional group identifier, and the random intercept is given by network identifiers nested within the study identifiers. To fit the LMMs models in our analyses, we used the ‘lmer’ package [@bates2015].

Once we obtained $p_i^O$ and $p_i^E$, we used simulation to determine whether the former is likely to come from the latter or not, since the large number of possible motif combinations and the small probabilities for some of them advise against using an Exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contains 10 million elements and, for each combination, the probability of being selected is equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. Following the usual interpretation for z-scores, motif combinations with $z_i^O > 1.96$ are over-represented, and those with $z_i^O < -1.96$ are under-represented, at the 95%  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).
